---
title: "Ocean Object Detection"
output: html_document
date: "2024-08-29"
---

# Module 1: ROI Definition and Imagery Acquisition
Define the region of interest (ROI), setup data access to Copernicus, and download the relevant imagery.
Package installation requirements:
1. rstac <- Client library for spatio-temporal asset catalog
2. sf <- simple features for R
3. stars <- spatio-temporal arrays, raster, and vector data cubes
4. terra <- Spatial data analysis
5. abind <- Combine multidimensional arrays
6. sp <- Classes and methods for spatial data

### Install packages
```{r}
install.packages ("rstac")
install.packages ("sf")
install.packages ("stars")
install.packages ("terra")
install.packages ("abind")
install.packages ("sp")
```

### Load libraries 
```{r}
library(rstac)
library(sf)
library(stars)
library(terra) 
library(sp)
library(abind)
```

## User Input
Define the key inputs for the module here:
1. Region of interest (ROI) by defining minimum/maximum decimal latitude/longitude.
2. Start and end time in the format: yyyy-mm-dd
3. Data output location.

### Define AOI
```{r}
bbox <- st_bbox(c(xmin=-0.4005585, ymin=5.422245, 
                 xmax=-0.1936628, ymax=5.544341),
                 crs=4326)
```

### Define start and end date of data acqusition
```{r}
start_date <- "2018-10-31" 
end_date <- "2018-11-01"
```


### Access the Sentinel-2 Database
```{r}
url_dataspace <-"https://planetarycomputer.microsoft.com/api/stac/v1/"
level = "sentinel-2-l2a"
```


### The query command to obtain the data
```{r}
matches <- 
  stac(url_dataspace) |>
  stac_search(collections = level,
              bbox = bbox,
              datetime = "2018-10-31/2018-11-01") |>
  get_request() |>
  items_sign(sign_fn = sign_planetary_computer())
best_img <- matches$features[[1]]
```

### AOI to crop the imagery
```{r}
red <- read_stars( paste0("/vsicurl/", best_img$assets$B04$href) )
bbox_proj <- bbox |> st_as_sfc() |> st_transform(st_crs(red)) |> vect()
```


### Download Image at ROI
```{r}
Blue <- rast( paste0("/vsicurl/", best_img$assets$B02$href) )|> crop(bbox_proj)
Green <- rast( paste0("/vsicurl/", best_img$assets$B03$href) )|> crop(bbox_proj)
Red <- rast( paste0("/vsicurl/", best_img$assets$B04$href) )|> crop(bbox_proj)
Rededge <- rast( paste0("/vsicurl/", best_img$assets$B06$href) )|> crop(bbox_proj)
NIR  <- rast( paste0("/vsicurl/", best_img$assets$B08$href) )|> crop(bbox_proj)
SWIR <- rast( paste0("/vsicurl/", best_img$assets$B11$href) )|> crop(bbox_proj)
```


### Resample Rededge2 and SWIR 2 from 20m to 10m**
```{r}
Rededge2 <- resample(Rededge, NIR, method='bilinear')
SWIR2 <- resample(SWIR, NIR, method='bilinear')
```

### Display True Colour Image at ROI
The true colour image that covers the ROI is displayed here, for Sentinel-2 level 2A data.

```{r}
raster_stack <- c(Red, Green, Blue)
raster_stack_all <- c(Blue, Green, Red, Rededge2, NIR, SWIR2)
plotRGB(raster_stack, scale = 1, stretch = "lin")
```

### Save Dataset
Save the multispectral Sentinel-2 data as a GeoTIFF

```{r}
writeRaster(raster_stack_all, "./Accra.tif", overwrite=TRUE)
```

# Module 2: Read Data and Detect Objects
Read GeoTIFF data and perform object detection using K-Means clustering.

### K-Means Clustering (NDVI)
```{r}
#NDVI
NDVI <-(NIR-Red)/(NIR+Red)
NDVI  |> plot()

#Classification
nr <- as.data.frame(NDVI, cell=TRUE)
set.seed(99)
kmncluster <- kmeans(nr[,-1], centers=2, iter.max = 500, nstart = 3, algorithm="Lloyd")
MOI_NDVI <- NDVI
MOI_NDVI[] <- kmncluster$cluster
plot(MOI_NDVI, main = 'Unsupervised classification of NDVI')
```

### K-Means Clustering (FDI)
```{r}
#FDI
FDI <- NIR-(Rededge2+(SWIR2-Rededge2)*((842-665)/(1610-665))*10)
FDI  |> plot()

#Classification
nr <- as.data.frame(FDI, cell=TRUE)
set.seed(99)
kmncluster <- kmeans(nr[,-1], centers=3, iter.max = 500, nstart = 3, algorithm="Lloyd")
MOI_FDI <- FDI
MOI_FDI[] <- kmncluster$cluster
plot(MOI_FDI, main = 'Unsupervised classification of FDI')
```

### Masking Land
```{r}
sf_object <- st_read("Sea_accra.shp")
spat_vector <- vect(sf_object)
shapefile_reprojected <- project(spat_vector, "EPSG:32630")
 
#check raster and vector is overlay
plot(NDVI)
plot(shapefile_reprojected, add = TRUE, border = "blue", lwd = 2)
 
#masking NDVI
mask_NDVI<- crop(MOI_NDVI, shapefile_reprojected)
mask_MOI_NDVI<- mask(mask_NDVI, shapefile_reprojected)
plot(mask_MOI_NDVI)

#masking FDI
mask_FDI<- crop(MOI_FDI, shapefile_reprojected)
mask_MOI_FDI<- mask(mask_FDI, shapefile_reprojected)
plot(mask_MOI_FDI)
```

# Module 3: Visualitation & Create Shapefile from Classified Areas

Export classified areas as shapefiles.

### Oil Spill Map based on NDVI
```{r}
plot(mask_MOI_NDVI, main = 'Oil Spill (NDVI) Map', legend = F,
     col = c('red', 'white'),
     )
legend('left', legend = c('Non oil spill', 'Oil spill'),
       fill = c('red', 'white'),
       border = FALSE,
       bty = 'n',
       xpd = T,horiz = F,
       inset = c(0.99, 1))
writeRaster(mask_MOI_NDVI, "./MOI_NDVI.tif", overwrite=TRUE)
```
### Oil Spill Map based on FDI
```{r}
plot(mask_MOI_FDI, main = 'Oil Spill (FDI) Map ', legend = F,
     col = c('white','red', 'white'),
     )
legend('left', legend = c( 'Land','Oil spill','Non oil spill'),
       fill = c('white','red', 'white'),
       border = FALSE,
       bty = 'n',
       xpd = T,horiz = F,
       inset = c(0.99, 1))
writeRaster(mask_MOI_FDI, "./MOI_FDI.tif", overwrite=TRUE)
```
### Convert raster to polygon
```{r}
#NDVI
polygon_Oilspill_NDVI <- as.polygons(mask_MOI_NDVI, dissolve = TRUE)
plot(polygon_Oilspill_NDVI, main = "Raster to Polygon (NDVI)")

#FDI
polygon_Oilspill_FDI <- as.polygons(mask_MOI_FDI, dissolve = TRUE)
plot(polygon_Oilspill_FDI, main = "Raster to Polygon (FDI)")
```